\documentclass[12pt,a4paper]{article}
\input{default_preamble.tex}
\input{homework_commands.tex}
\input{default_gossary_entries.tex}

\usepackage{fontspec}
\usepackage{minted}

\begin{document}

% \author{Student}
\title{Computational homework}
\date{\today}
\maketitle

%%--CABEÇALHO--%%
\begin{center}
    {\huge Simulink-based implementation of an \(M\)-QAM modem \par}
    {\LARGE Digital communication systems \par}
    {\Large Telecommunication Engineering \par}
    {\Large André Lima Ferrer de Almeida \par}
    % {\Large Opitional}
\end{center}

\problem \label{probone}
Develop a Matlab function that calculates a bit sequence in Gray encoding. It must input the number of bits (an integer value) and output the bit sequence. For example, for a sequence of \(3\) bits, your function must return

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\small, framesep=2mm]{matlab}
    >> get_gray_codes(3)

ans =

     0     1     1     0     0     1     1     0
     0     0     1     1     1     1     0     0
     0     0     0     0     1     1     1     1
\end{minted}
where each column corresponds to a sequence (from the left to the right) of 3 bits in Gray encoding. The first row is the \gls{lsb} and the last row is the \gls{msb}.

Tips:
\begin{enumerate}
    \item Before generating the Gray encoding, first generate the bit sequence without any encoding, i.e.,
    \begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\small, framesep=2mm]{matlab}
        >> original

    original =

     0     1     0     1     0     1     0     1
     0     0     1     1     0     0     1     1
     0     0     0     0     1     1     1     1
    \end{minted}
    \item Transforming the uncoded bit sequence into Gray encoding is easier if you use a recursive function. \href{https://en.wikipedia.org/wiki/Gray_code#Constructing_an_n-bit_Gray_code}{See here}.
\end{enumerate}

\textbf{Note:} For the following problems, use the Simulink template provided by the professor.

\section{Baseband modulator}

\problem \label{probtwo}
Generate a bit stream with the following parameters:
\begin{itemize}
    \item Bit rate: \(R_b = \qty{10}{\mega\bit\per\second}\).
    \item Equiprobable bits.
\end{itemize}

\problem \label{probthree}
Use a block called ``Matlab function'' to create a function that maps a stream of \(N_b = \log_2\left( M \right)\) bits to a complex symbol from a constellation of \(M=256\) symbols. The real and imaginary part of each constellation point, must be spaced as Proakis suggests, i.e., \(\left\{ \pm1, \pm3, \dots, \pm\sqrt{M} \right\}\). Moreover, the mapping of the real/imaginary part to the complex symbol must be in the Gray encoding sequence (see \cite{proakisDigitalCommunications2007} for more info).

\subproblem What is the new operating rate of the system?
\subproblem Use a block called ``Gain'' to apply a gain of \(\sqrt{E_g}\) to the complex symbol outputs, where \(E_g = 9\).
\subproblem By using a block called ``Constellation Diagram'', show that each point in the constellation is now in the form \(\sqrt{E_g} A_m\), where \(A_m = A_{mi}+j A_{mq}\) is the constellation for the \(m\)th symbol and that \(A_{mi}, A_{mq} \in \left\{ \pm1, \pm3, \dots, \pm\sqrt{M} \right\}\). Note that the \(k\)th transmitted symbol can be represented by the vector
\begin{align}
    \mathbf{s}_m\left[ k \right] = \begin{bmatrix}
        A_{mi} & A_{mq}
    \end{bmatrix}^\top \in \mathbb{R}^{2}
\end{align}

Tips:
\begin{enumerate}
    \item Use the ``Buffer'' block to vectorize the bit stream to \(N_b\) symbols and use it as input of the ``Matlab function''.
    \item You can pass a matrix of the sequence of bits with Gray encoding to the ``Matlab function'' block. You can use your function developed in Problem \ref{probone} to generate this matrix.
\end{enumerate}

\problem Pass your symbol stream through the ``Raised Cosine Filter block'' with the following parameters
\begin{itemize}
    \item Roll-off factor: \(0.2\).
    \item Filter span (in symbols): \(4\).
    \item Number of samples per symbol: \(N=100\) samples.
\end{itemize}
The filter gain must be adjusted so that the energy of the discrete-time impulse response is \(1\).

Tips:
\begin{enumerate}
    \item You can obtain the impulse response of the Raised Cosine Filter and compute numerically its energy (it is an \gls{fir} filter as the impulse response is truncated).
    \item Use the ``square'' and the ``mean'' block to check if the input and output energy remain unchanged after the filtering.
\end{enumerate}
\subproblem What is the new operating rate of the system? Why has the operating rate changed to this value?
\subproblem Check if the raised cosine filter obeys the Nyquist criterion for no \gls{isi}. Use the ``scope'' block to prove your answer graphically.

\section{Upconverter}

\problem Use a block called ``sine wave'' (from the DSP system toolbox) with the following parameters
\begin{itemize}
    \item Amplitude: \(\sqrt{2}\);
    \item Frequency: \(f_c = \qty{100}{\mega\hertz}\);
    \item Sample mode: Discrete;
    \item Output complexity: Complex;
\end{itemize}

The sampling rate must be consistent with the sampling rate of your system. Let us denote the \gls{bb} pulse shaping (i.e., the raised cosine filter) of your system as \(g\left( t \right) \in \mathbb{R}\), whose energy is denoted as \(E_g\). Then the pre-envelope (or analytic signal) of the \(m\)th symbol is
\begin{align}
    s_{m+}\left( t \right) = s_{ml}\left( t \right) e^{j 2\pi f_c t} \in \mathbb{C},
\end{align}
where \(s_{ml}\left( t \right) = A_m g\left( t \right) \in \mathbb{C}\) is the complex envelope (or lowpass equivalent) of the real-valued transmitted signal for the \(m\)th symbol,
\begin{align}
    s_{m}\left( t \right) = \sqrt{2} \Re{s_{ml}\left( t \right) e^{j 2\pi f_c t}}.
\end{align}
The pre-envelope has important properties concerning the spectrum of the transmitted signal. The output of the block named ``sine wave'' should exactly be the pre-envelope discussed in the literature. By using the ``Spectrum Analyzer'' block, make a mathematical analysis of this signal, and compare its theoretical spectrum with the spectrum obtained in Simulink. Good references can be found in \cite{mengaliSynchronizationTechniquesDigital1997,meyrDigitalCommunicationReceivers1998,proakisDigitalCommunications2007}

%============================================
\printbibliography


\end{document}